Structured Versus Decorated Cospans

### 2020-11-09 cospans.tex
lots of small improvements

\author{John\ C.\ Baez$^{*,\bullet}$, Kenny Courser$^*$, and Christina Vasilakopoulou$^\square$}
\address{$^*$Department of Mathematics, University of California, Riverside CA, USA 92521}
\address{$^\bullet$Centre for Quantum Technologies, National University of Singapore, Singapore 117543}
\address{$^\square$Department of Mathematics, University of Patras, Greece 265 04}
\email{baez@ucr.edu, kcour001@ucr.edu, cvasilak@math.upatrs.gr}

@@

Our approach is to instead view the functor $F$ as a pseudofunctor $F \maps \textrm{A} \to \mathbf{Cat}$ and take advantage of the 2-categorical structure of $\mathbf{Cat}$. Given an object $a \in \mathsf{A}$, this then allows us to view $F(a)$ not as a \emph{set} of decorations on the object $a$ but as a \emph{category} of decorations on the object $a$. From this pseudofunctor, we construct a symmetric monoidal double category $F \mathbb{C}$sp. Double categories were first introduced by Ehresmann \cite{Ehresmann63, Ehresmann65} and symmetric monoidal double categories by Shulman \cite{Shul}. They have long been used in topology and other branches of pure mathematics \cite{Brown1,Brown2}.  More recently they have been used to study open dynamical systems \cite{LS} and open Markov processes \cite{BC}. While a mere category has only objects and morphisms, a double category has a few more types of entities:
\[
\begin{tikzpicture}[scale=1]
\node (D) at (-4,0.5) {$A$};
\node (E) at (-2,0.5) {$B$};
\node (F) at (-4,-1) {$C$};
\node (A) at (-2,-1) {$D$};
\node (B) at (-3,-0.25) {$\Downarrow a$};
\path[->,font=\scriptsize,>=angle 90]
(D) edge node [above]{$M$}(E)
(E) edge node [right]{$g$}(A)
(D) edge node [left]{$f$}(F)
(F) edge node [above]{$N$} (A);
\end{tikzpicture}
\]
We call $A, B, C$ and $D$ `objects', $f$ and $g$ `vertical 1-morphisms', $M$ and $N$ `horizontal 1-cells', and $a$ a `2-morphism'.   We can compose vertical 1-morphisms to get new vertical 1-morphisms and compose horizontal 1-cells to get new horizontal 1-cells.  We can compose the 2-morphisms in two ways: horizontally by setting squares side by side, and vertically by setting one on top of the other.   In a `strict' double category all these forms of composition are associative.  In a `pseudo' double category, horizontal 1-cells compose in a weakly associative manner: that is, the associative law holds only up to an invertible 2-morphism, called the `associator', which obeys a coherence law.

@@

\section{A symmetric monoidal double category of decorated cospans}\label{DecCospansDoubleCat}
In this section we build the symmetric monoidal double category $F \mathbb{C}$sp mentioned in the introduction. Then we obtain an underlying symmetric monoidal bicategory $F \mathbf{Csp}$ using a result of Shulman \cite{Shul} and then finally decategorify the obtained symmetric monoidal bicategory to obtain a symmetric monoidal category $F$Csp which is a generalization of Fong's original decorated cospan category \cite{Fong}. 
The definition of a lax monoidal pseudofunctor is recalled in \cref{sec:2cats}, and the basics of double categories are recalled in \cref{sec:doublecats}.

@@

Let $\mathsf{A}$ be a category with finite colimits and $F \maps \mathsf{A} \to \mathbf{Cat}$ a lax monoidal pseudofunctor. Then there exists a (pseudo) double category $F\mathbb{C}$sp which has:

@@

where $!_a \in F(a)$ is the trivial decoration on $a$ given by the composite of the unique map $F(!) \maps F(0) \to F(a)$ and the morphism $\phi \maps 1 \to F(0)$  which comes from the structure of the lax monoidal pseudofunctor $F \maps \mathsf{A} \to \mathbf{Cat}$. For morphisms, the structure functor $U$ is defined as:

@@

Given two composable horizontal 1-cells $M$ and $N$:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1,0) {$m$};
\node (C) at (2,0) {$b$};
\node (D) at (1,-0.5) {$x \in F(m)$};
\node (E) at (3,0) {$b$};
\node (F) at (4,0) {$n$};
\node (G) at (5,0) {$c$};
\node (H) at (4,-0.5) {$y \in F(n)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B)
(E) edge node[above]{$i^\prime$} (F)
(G) edge node[above]{$o^\prime$} (F);
\end{tikzpicture}
\]
the composite $N \odot M$ is given by:
\begin{equation}\label{eq:dcospanscomposition}
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1,1) {$m$};
\node (C) at (2,0) {$b$};
\node (D) at (3,1) {$n$};
\node (E) at (4,0) {$c$};
\node (F) at (2,2) {$m+n$};
\node (G) at (2,3) {$m+_{b} n$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B)
(C) edge node [above] {$i^\prime$} (D)
(E) edge node [above] {$o^\prime$} (D)
(B) edge node [above] {$j$} (F)
(D) edge node [above] {$j^\prime$} (F)
(F) edge node [left] {$\psi$} (G)
(A) edge[bend left] node [left] {$\psi j i$} (G)
(E) edge[bend right] node [right] {$\psi j^\prime o^\prime$} (G);
\end{tikzpicture}
\end{equation}
with the corresponding decoration of the apex $x \odot y \in F(m+_{b} n)$ given by:
$$1 \xrightarrow{\lambda^{-1}} 1 \times 1 \xrightarrow{x \times y} F(m) \times F(n) \xrightarrow{\phi_{m,n}} F(m+n) \xrightarrow{F(\psi)} F(m+_{b}n)$$
where $\psi \maps m + n \to m+_{b} n$ is the natural map from the coproduct to the pushout and $\phi_{m,n} \maps F(m) \times F(n) \to F(m+n)$ is the natural transformation coming from the structure of the lax monoidal pseudofunctor $F \maps \mathsf{A} \to \mathbf{Cat}$. Denoting the first and second of these horizontal 1-cells as $M$ and $N$, respectively, the source and target structure functors satisfy the equations $S(N \odot M)=S(M)$ and $T(N \odot M)=T(N)$.

@@

These two 2-morphisms $\iota_\alpha$ and $\iota_\beta$ are two 2-morphisms in the monoidal 2-category $(\mathbf{Cat},\times,1)$ and so we can tensor them which results in:

@@

As is usual concerning the interchange law of double categories of this nature, only the `interior' of the two composites appears different, but the two morphisms $(h_1^\prime +_{g^\prime} h_2^\prime)(h_1 +_g h_2) \maps m+_{b} n \to m'' +_{b''} n''$ and $(h_1^\prime h_1) +_{g^\prime g} (h_2^\prime h_2) \maps m+_{b} n \to m'' +_{b''}n''$ are the same universal map realized in two different ways. The two morphisms of decorations $\iota_{(\alpha^\prime \odot \beta^\prime)(\alpha \odot \beta)}$ and $\iota_{(\alpha^\prime \alpha) \odot (\beta^\prime \beta)}$ are obtained as two different compositions of four 2-morphisms in $\mathbf{Cat}$, namely horizontally then vertically and vertically then horizontally. As $\mathbf{Cat}$ is a 2-category, the interchange law for these 2-morphisms already holds, and as a result, the morphisms $$\iota_{(\alpha' \odot \beta')(\alpha \odot \beta)} \maps F((h_1^\prime +_{g^\prime} h_2^\prime)(h_1 +_g h_2))(x \odot y) \to x'' \odot y''$$ and $$\iota_{(\alpha^\prime \alpha)\odot(\beta^\prime \beta)} \maps F((h_1^\prime h_1)+_{g^\prime g} (h_2^\prime h_2))(x \odot y) \to x'' \odot y''$$ are also the same. Thus the interchange law for 2-morphisms holds and $F\mathbb{C}$sp is a double category.

@@

Let $\mathsf{A}$ be a category with finite colimits and $F \maps \mathsf{A} \to \mathbf{Cat}$ a symmetric lax monoidal pseudofunctor. Then the double category $F\mathbb{C}$sp is symmetric monoidal.

@@

where $0$ is the monoidal unit of $\mathsf{A}$ and $!_0 \maps 1 \to F(0)$ is the morphism which is part of the structure of the symmetric lax monoidal pseudofunctor $F \maps \mathsf{A} \to \mathbf{Cat}$. Tensoring an object with the monoidal unit, say, on the left:

@@

since $$F(aa)(((x_1+x_2)+x_3)+x_4)=F((1 \otimes a)a(a \otimes 1))(((x_1+x_2)+x_3)+x_4)$$ as the corresponding pentagon in the symmetric monoidal category $(\mathsf{A},+,0)$ commutes, and then appliying the pseudofunctor $F$ to said pentagon yields a commutative pentagon in $\mathbf{Cat}$.

Similarly, if we denote the left and right unitors as $\lambda$ and $\rho$, respectively, then the following triangle of cospans and underlying maps of cospans commutes:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$(M_1 \otimes 0) \otimes M_2$};
\node (B) at (2.25,1) {$M_1 \otimes M_2$};
\node (C) at (4.5,0) {$M_1 \otimes (0 \otimes M_2)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$\rho \otimes 1$} (B)
(C) edge node[above]{$1 \otimes \lambda$} (B)
(A) edge node[above]{$a$} (C);
\end{tikzpicture}
\]
as well as the following triangle of corresponding decorations in the category $F(m_1+m_2)$:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$F(\rho \otimes 1)((x_1+0)+x_2)$};
\node (B) at (2.25,1) {$x_1+x_2$};
\node (C) at (4.5,0) {$F(1 \otimes \lambda)(x_1+(0+x_2)) $};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[left]{$\iota_{\rho \otimes 1}$} (B)
(C) edge node[right]{$\iota_{1 \otimes \lambda}$} (B)
(A) edge node[above]{$F(1 \otimes \lambda)(\iota_a)$} (C);
\end{tikzpicture}
\]
since $$F(\rho \otimes 1)((x_1+0)+x_2)=F((1 \otimes \lambda)a)((x_1+0)+x_2)$$ as the corresponding triangle in the symmetric monoidal category $(\mathsf{A},+,0)$ commutes and applying the pseudofunctor $F$ to this commutative triangle results in a commutative triangle in $\mathbf{Cat}$.

@@

as the above underlying diagram of maps of cospans commutes and then applying the pseudofunctor $F$ yields a commutative diagram in $\mathbf{Cat}$.

@@

as the above underlying diagram of maps of cospans commutes and then applying the pseudofunctor $F$ to this diagram yields a commutative diagram in $\mathbf{Cat}$. The other diagrams are shown to commute similarly.

@@

\begin{thm}[Shulman]\label{Shul}
Let $\mathbb{X}$ be an isofibrant symmetric monoidal pseudo double category. Then the horizontal bicategory $H(\mathbb{X})$ of $\mathbb{X}$ is a symmetric monoidal bicategory which has:
\begin{enumerate}
\item{objects as those of $\mathbb{X}$,}
\item{morphisms as horizontal 1-cells of $\mathbb{X}$, and}
\item{2-morphisms as globular 2-morphisms of $\mathbb{X}$.}
\end{enumerate}
\end{thm}

\begin{cor}
Let $(\mathsf{C},+,0)$ be a category with finite colimts and $F \maps \mathsf{C} \to \mathbf{Cat}$ a symmetric lax monoidal pseudofunctor. Then there exists a symmetric monoidal bicategory $F \mathbf{Csp} \mapseqq H(F\mathbb{C}\textnormal{sp})$ which has:
\begin{enumerate}
\item{objects as those of $\mathsf{A}$,}
\item{morphisms as $F$-decorated cospans:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1,0) {$c$};
\node (C) at (2,0) {$b$};
\node (D) at (4,0) {$d \in F(c)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B);
\end{tikzpicture}
\]
and}
\item{2-morphisms as maps of cospans in $\mathsf{A}$ of the form:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1,0.5) {$c$};
\node (C) at (2,0) {$b$};
\node (E) at (1,-0.5) {$c^\prime$};
\node (D) at (3,0.5) {$d \in F(c)$};
\node (F) at (3,-0.5) {$d^\prime \in F(c^\prime)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B)
(A) edge node[below]{$i^\prime$} (E)
(B) edge node[left]{$h$} (E)
(C) edge node[below]{$o^\prime$} (E);
\end{tikzpicture}
\]
together with a morphism $\iota \maps F(h)(d) \to d^\prime$ in $F(c^\prime)$.}
\end{enumerate}
\end{cor}

\begin{proof}
This follows immediately by Shulman's Theorem \ref{Shul} above applied to the fibrant symmetric monoidal double category $F\mathbb{C}\textnormal{sp}$.
\end{proof}

@@

\begin{cor}
Given a symmetric lax monoidal pseudofunctor $F \maps \mathsf{A} \to \mathbf{Cat}$ where $\mathsf{A}$ is a category with finite colimits and whose monoidal structure is given by binary coproducts, there exists a symmetric monoidal category $F\textnormal{Csp} \mapseqq D(F\mathbf{Csp})$ which has:
\begin{enumerate}
\item{objects as those of $\mathsf{A}$ and}
\item{morphisms as isomorphism classes of $F$-decorated cospans of $\mathsf{A}$, where an $F$-decorated cospan is given by a pair:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1,0) {$c$};
\node (C) at (2,0) {$b$};
\node (D) at (4,0) {$d \in F(c)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B);
\end{tikzpicture}
\]
Given another $F$-decorated cospan:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1,0) {$c^\prime$};
\node (C) at (2,0) {$b$};
\node (D) at (4,0) {$d^\prime \in F(c^\prime)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[below]{$i^\prime$} (B)
(C) edge node[below]{$o^\prime$} (B);
\end{tikzpicture}
\]
these two $F$-decorated cospans are in the same isomorphism class if there exists an isomorphism $f \maps c \to c^\prime$ such that following diagram commutes:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B') at (1,0.5) {$c$};
\node (B) at (1,-0.5) {$c^\prime$};
\node (C) at (2,0) {$b$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[below]{$i^\prime$} (B)
(C) edge node[below]{$o^\prime$} (B)
(A) edge node[above]{$i$} (B')
(C) edge node[above]{$o$} (B')
(B') edge node[left]{$f$} (B);
\end{tikzpicture}
\]
and there exists an isomorphism $\iota \maps F(f)(d) \to d^\prime$ in $F(c^\prime)$.}
\end{enumerate}
\end{cor}
In this symmetric monoidal category, isomorphism classes are as they should morally be, and the instance of two graphs having different edge sets does not prevent them from being in the same isomorphism class due to the isomorphism $\iota$.

%Let $F \maps \mathsf{A} \to \mathbf{Cat}$ be a symmetric lax monoidal pseudofunctor. Then by Theorem \ref{DC} of the previous section, we get a symmetric monoidal double category $F\mathbb{C}\textnormal{sp}$. This symmetric monoidal double category has:
%\begin{enumerate}
%\item{objects as those of $\mathsf{A}$,}
%\item{vertical 1-morphisms as morphisms of $\mathsf{A}$,}
%\item{horizontal 1-cells as pairs:
%\[
%\begin{tikzpicture}[scale=1.5]
%\node (A) at (0,0) {$a$};
%\node (B) at (1,0) {$c$};
%\node (C) at (2,0) {$b$};
%\node (D) at (4,0) {$d \in F(c)$};
%\path[->,font=\scriptsize,>=angle 90]
%(A) edge node[above]{$i$} (B)
%(C) edge node[above]{$o$} (B);
%\end{tikzpicture}
%\]
%and}
%\item{2-morphisms as maps of cospans in $\mathsf{A}$
%\[
%\begin{tikzpicture}[scale=1.5]
%\node (A) at (0,0.5) {$a$};
%\node (A') at (0,-0.5) {$a^\prime$};
%\node (B) at (1,0.5) {$c$};
%\node (C) at (2,0.5) {$b$};
%\node (C') at (2,-0.5) {$b^\prime$};
%\node (D) at (1,-0.5) {$c^\prime$};
%\node (E) at (3,0.5) {$d \in F(c)$};
%\node (F) at (3,-0.5) {$d^\prime \in F(c^\prime)$};
%\path[->,font=\scriptsize,>=angle 90]
%(A) edge node[above]{$$} (B)
%(C) edge node[above]{$$} (B)
%(A) edge node[left]{$f$} (A')
%(C) edge node[left]{$g$} (C')
%(A') edge node {$$} (D)
%(C') edge node {$$} (D)
%(B) edge node [left] {$h$} (D);
%\end{tikzpicture}
%\]
%together with a morphism $\iota \maps F(h)(d) \to d^\prime$ in $F(c^\prime)$.}
%\end{enumerate}

>>>>

\author{John\ C.\ Baez$^{1,2}$, Kenny Courser$^1$, and Christina Vasilakopoulou$^3$}
\address{$^1$Department of Mathematics, University of California, Riverside CA, USA 92521}
\address{$^2$Centre for Quantum Technologies, National University of Singapore, Singapore 117543}
\address{$^3$Department of Mathematics, University of Patras, Greece 265 04}
\email{baez@math.ucr.edu, kcour001@ucr.edu, cvasilak@math.upatrs.gr}

@@

Our approach is to instead view the functor $F$ as a pseudofunctor $F \maps \textrm{A} \to \Cat$ and take advantage of the 2-categorical structure of $\Cat$. Given an object $a \in \mathsf{A}$, this then allows us to view $F(a)$ not as a \emph{set} of decorations on the object $a$ but as a \emph{category} of decorations on the object $a$. From this pseudofunctor, we construct a symmetric monoidal double category $F \mathbb{C}$sp. Double categories were first introduced by Ehresmann \cite{Ehresmann63, Ehresmann65} and symmetric monoidal double categories by Shulman \cite{Shul}. They have long been used in topology and other branches of pure mathematics \cite{Brown1,Brown2}.  More recently they have been used to study open dynamical systems \cite{LS} and open Markov processes \cite{BC}. While a mere category has only objects and morphisms, a double category has a few more types of entities:
\[
\begin{tikzpicture}[scale=1]
\node (D) at (-4,0.5) {$A$};
\node (E) at (-2,0.5) {$B$};
\node (F) at (-4,-1) {$C$};
\node (A) at (-2,-1) {$D$};
\node (B) at (-3,-0.25) {$\Downarrow a$};
\path[->,font=\scriptsize,>=angle 90]
(D) edge node [above]{$M$}(E)
(E) edge node [right]{$g$}(A)
(D) edge node [left]{$f$}(F)
(F) edge node [above]{$N$} (A);
\end{tikzpicture}
\]
We call $A, B, C$ and $D$ `objects', $f$ and $g$ `vertical 1-morphisms', $M$ and $N$ `horizontal 1-cells', and $a$ a `2-morphism'.   We can compose vertical 1-morphisms to get new vertical 1-morphisms and compose horizontal 1-cells to get new horizontal 1-cells.  We can compose the 2-morphisms in two ways: horizontally by setting squares side by side, and vertically by setting one on top of the other.   In a `strict' double category all these forms of composition are associative.  In a `pseudo' double category, horizontal 1-cells compose in a weakly associative manner: that is, the associative law holds only up to an invertible 2-morphism, called the `associator', which obeys a coherence law.

@@

\section{A symmetric monoidal double category of decorated cospans}\label{DecCospansDoubleCat}
In this section we build the symmetric monoidal double category $F \mathbb{C}$sp mentioned in the introduction. Then we obtain an underlying symmetric monoidal bicategory $F \mathbf{Csp}$ using a result of Shulman \cite{Shul} and finally decategorify this to obtain a symmetric monoidal category $F\mathsf{Csp}$ which is a generalization of Fong's original decorated cospan category \cite{Fong}.  The definition of a lax monoidal pseudofunctor is recalled in \cref{sec:2cats}, and the basics of double categories are recalled in \cref{sec:doublecats}.

@@

Let $\mathsf{A}$ be a category with finite colimits and $F \maps \mathsf{A} \to \Cat$ a lax monoidal pseudofunctor. Then there exists a (pseudo) double category $F\mathbb{C}$sp which has:

@@

where $!_a \in F(a)$ is the trivial decoration on $a$ given by the composite of the unique map $F(!) \maps F(0) \to F(a)$ and the morphism $\phi \maps 1 \to F(0)$  which comes from the structure of the lax monoidal pseudofunctor $F \maps \mathsf{A} \to \Cat$. For morphisms, the structure functor $U$ is defined as:

@@

Given two composable horizontal 1-cells $M$ and $N$:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1,0) {$m$};
\node (C) at (2,0) {$b$};
\node (D) at (1,-0.5) {$x \in F(m)$};
\node (E) at (3,0) {$b$};
\node (F) at (4,0) {$n$};
\node (G) at (5,0) {$c$};
\node (H) at (4,-0.5) {$y \in F(n)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B)
(E) edge node[above]{$i^\prime$} (F)
(G) edge node[above]{$o^\prime$} (F);
\end{tikzpicture}
\]
the composite $N \odot M$ is given by:
\begin{equation}\label{eq:dcospanscomposition}
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1,1) {$m$};
\node (C) at (2,0) {$b$};
\node (D) at (3,1) {$n$};
\node (E) at (4,0) {$c$};
\node (F) at (2,2) {$m+n$};
\node (G) at (2,3) {$m+_{b} n$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B)
(C) edge node [above] {$i^\prime$} (D)
(E) edge node [above] {$o^\prime$} (D)
(B) edge node [above] {$j$} (F)
(D) edge node [above] {$j^\prime$} (F)
(F) edge node [left] {$\psi$} (G)
(A) edge[bend left] node [left] {$\psi j i$} (G)
(E) edge[bend right] node [right] {$\psi j^\prime o^\prime$} (G);
\end{tikzpicture}
\end{equation}
with the corresponding decoration of the apex $x \odot y \in F(m+_{b} n)$ given by:
$$1 \xrightarrow{\lambda^{-1}} 1 \times 1 \xrightarrow{x \times y} F(m) \times F(n) \xrightarrow{\phi_{m,n}} F(m+n) \xrightarrow{F(\psi)} F(m+_{b}n)$$
where $\psi \maps m + n \to m+_{b} n$ is the natural map from the coproduct to the pushout and $\phi_{m,n} \maps F(m) \times F(n) \to F(m+n)$ is the natural transformation coming from the structure of the lax monoidal pseudofunctor $F \maps \mathsf{A} \to \Cat$. Denoting the first and second of these horizontal 1-cells as $M$ and $N$, respectively, the source and target structure functors satisfy the equations $S(N \odot M)=S(M)$ and $T(N \odot M)=T(N)$.

@@

These two 2-morphisms $\iota_\alpha$ and $\iota_\beta$ are two 2-morphisms in the monoidal 2-category $(\Cat,\times,1)$ and so we can tensor them which results in:

@@

As is usual concerning the interchange law of double categories of this nature, only the `interior' of the two composites appears different, but the two morphisms $(h_1^\prime +_{g^\prime} h_2^\prime)(h_1 +_g h_2) \maps m+_{b} n \to m'' +_{b''} n''$ and $(h_1^\prime h_1) +_{g^\prime g} (h_2^\prime h_2) \maps m+_{b} n \to m'' +_{b''}n''$ are the same universal map realized in two different ways. The two morphisms of decorations $\iota_{(\alpha^\prime \odot \beta^\prime)(\alpha \odot \beta)}$ and $\iota_{(\alpha^\prime \alpha) \odot (\beta^\prime \beta)}$ are obtained as two different compositions of four 2-morphisms in $\Cat$, namely horizontally then vertically and vertically then horizontally. As $\Cat$ is a 2-category, the interchange law for these 2-morphisms already holds, and as a result, the morphisms $$\iota_{(\alpha' \odot \beta')(\alpha \odot \beta)} \maps F((h_1^\prime +_{g^\prime} h_2^\prime)(h_1 +_g h_2))(x \odot y) \to x'' \odot y''$$ and $$\iota_{(\alpha^\prime \alpha)\odot(\beta^\prime \beta)} \maps F((h_1^\prime h_1)+_{g^\prime g} (h_2^\prime h_2))(x \odot y) \to x'' \odot y''$$ are also the same. Thus the interchange law for 2-morphisms holds and $F\mathbb{C}$sp is a double category.

@@

Let $\mathsf{A}$ be a category with finite colimits and $F \maps \mathsf{A} \to \Cat$ a symmetric lax monoidal pseudofunctor. Then the double category $F\mathbb{C}$sp is symmetric monoidal.

@@

where $0$ is the monoidal unit of $\mathsf{A}$ and $!_0 \maps 1 \to F(0)$ is the morphism which is part of the structure of the symmetric lax monoidal pseudofunctor $F \maps \mathsf{A} \to \Cat$. Tensoring an object with the monoidal unit, say, on the left:

@@

since $$F(aa)(((x_1+x_2)+x_3)+x_4)=F((1 \otimes a)a(a \otimes 1))(((x_1+x_2)+x_3)+x_4)$$ as the corresponding pentagon in the symmetric monoidal category $(\mathsf{A},+,0)$ commutes, and then appliying the pseudofunctor $F$ to said pentagon yields a commutative pentagon in $\Cat$.

Similarly, if we denote the left and right unitors as $\lambda$ and $\rho$, respectively, then the following triangle of cospans and underlying maps of cospans commutes:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$(M_1 \otimes 0) \otimes M_2$};
\node (B) at (2.25,1) {$M_1 \otimes M_2$};
\node (C) at (4.5,0) {$M_1 \otimes (0 \otimes M_2)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$\rho \otimes 1$} (B)
(C) edge node[above]{$1 \otimes \lambda$} (B)
(A) edge node[above]{$a$} (C);
\end{tikzpicture}
\]
as well as the following triangle of corresponding decorations in the category $F(m_1+m_2)$:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$F(\rho \otimes 1)((x_1+0)+x_2)$};
\node (B) at (2.25,1) {$x_1+x_2$};
\node (C) at (4.5,0) {$F(1 \otimes \lambda)(x_1+(0+x_2)) $};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[left]{$\iota_{\rho \otimes 1}$} (B)
(C) edge node[right]{$\iota_{1 \otimes \lambda}$} (B)
(A) edge node[above]{$F(1 \otimes \lambda)(\iota_a)$} (C);
\end{tikzpicture}
\]
since $$F(\rho \otimes 1)((x_1+0)+x_2)=F((1 \otimes \lambda)a)((x_1+0)+x_2)$$ as the corresponding triangle in the symmetric monoidal category $(\mathsf{A},+,0)$ commutes and applying the pseudofunctor $F$ to this commutative triangle results in a commutative triangle in $\Cat$.

@@

as the above underlying diagram of maps of cospans commutes and then applying the pseudofunctor $F$ yields a commutative diagram in $\Cat$.

@@

as the above underlying diagram of maps of cospans commutes and then applying the pseudofunctor $F$ to this diagram yields a commutative diagram in $\Cat$. The other diagrams are shown to commute similarly.

@@

\begin{lem}[Shulman]\label{Shul}
Let $\mathbb{X}$ be an isofibrant symmetric monoidal pseudo double category. Then the horizontal bicategory $H(\mathbb{X})$ of $\mathbb{X}$ is a symmetric monoidal bicategory which has:
\begin{enumerate}
\item{objects as those of $\mathbb{X}$,}
\item{morphisms as horizontal 1-cells of $\mathbb{X}$, and}
\item{2-morphisms as globular 2-morphisms of $\mathbb{X}$.}
\end{enumerate}
\end{lem}

\begin{thm}
Let $\A$ be a category with finite colimits and $F \maps (\A, +) \to \Cat$ a symmetric lax monoidal pseudofunctor. Then there exists a symmetric monoidal bicategory $F \mathbf{Csp} \mapseqq H(F\mathbb{C}\textnormal{sp})$ which has:
\begin{enumerate}
\item{objects as those of $\mathsf{A}$,}
\item{morphisms as $F$-decorated cospans:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1,0) {$c$};
\node (C) at (2,0) {$b$};
\node (D) at (4,0) {$d \in F(c)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B);
\end{tikzpicture}
\]
and}
\item{2-morphisms as maps of cospans in $\mathsf{A}$ of the form:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1,0.5) {$c$};
\node (C) at (2,0) {$b$};
\node (E) at (1,-0.5) {$c^\prime$};
\node (D) at (3,0.5) {$d \in F(c)$};
\node (F) at (3,-0.5) {$d^\prime \in F(c^\prime)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B)
(A) edge node[below]{$i^\prime$} (E)
(B) edge node[left]{$h$} (E)
(C) edge node[below]{$o^\prime$} (E);
\end{tikzpicture}
\]
together with a morphism $\iota \maps F(h)(d) \to d^\prime$ in $F(c^\prime)$.}
\end{enumerate}
\end{thm}

\begin{proof}
This follows immediately by applying \cref{Shul} to the fibrant symmetric monoidal double category $F\mathbb{C}\textnormal{sp}$.
\end{proof}

@@

\begin{cor}
Let $\A$ be a category with finite colimits and $F \maps (\A, +) \to \Cat$ a symmetric lax monoidal pseudofunctor.  Thenthere exists a symmetric monoidal category $F\textnormal{Csp}$ which has:
\begin{enumerate}
\item{objects as those of $\mathsf{A}$ and}
\item{morphisms as isomorphism classes of $F$-decorated cospans of $\mathsf{A}$, where an $F$-decorated cospan is given by a pair:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1,0) {$c$};
\node (C) at (2,0) {$b$};
\node (D) at (4,0) {$d \in F(c)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B);
\end{tikzpicture}
\]
Given another $F$-decorated cospan:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1,0) {$c^\prime$};
\node (C) at (2,0) {$b$};
\node (D) at (4,0) {$d^\prime \in F(c^\prime)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[below]{$i^\prime$} (B)
(C) edge node[below]{$o^\prime$} (B);
\end{tikzpicture}
\]
these two $F$-decorated cospans are in the same isomorphism class if there exists an isomorphism $f \maps c \to c^\prime$ such that following diagram commutes:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B') at (1,0.5) {$c$};
\node (B) at (1,-0.5) {$c^\prime$};
\node (C) at (2,0) {$b$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[below]{$i^\prime$} (B)
(C) edge node[below]{$o^\prime$} (B)
(A) edge node[above]{$i$} (B')
(C) edge node[above]{$o$} (B')
(B') edge node[left]{$f$} (B);
\end{tikzpicture}
\]
and there exists an isomorphism $\iota \maps F(f)(d) \to d^\prime$ in $F(c^\prime)$.}
\end{enumerate}
\end{cor}
In this symmetric monoidal category, isomorphism classes are as they should morally be, and the instance of two graphs having different edge sets does not prevent them from being in the same isomorphism class due to the isomorphism $\iota$.

%Let $F \maps \mathsf{A} \to \Cat$ be a symmetric lax monoidal pseudofunctor. Then by Theorem \ref{DC} of the previous section, we get a symmetric monoidal double category $F\mathbb{C}\textnormal{sp}$. This symmetric monoidal double category has:
%\begin{enumerate}
%\item{objects as those of $\mathsf{A}$,}
%\item{vertical 1-morphisms as morphisms of $\mathsf{A}$,}
%\item{horizontal 1-cells as pairs:
%\[
%\begin{tikzpicture}[scale=1.5]
%\node (A) at (0,0) {$a$};
%\node (B) at (1,0) {$c$};
%\node (C) at (2,0) {$b$};
%\node (D) at (4,0) {$d \in F(c)$};
%\path[->,font=\scriptsize,>=angle 90]
%(A) edge node[above]{$i$} (B)
%(C) edge node[above]{$o$} (B);
%\end{tikzpicture}
%\]
%and}
%\item{2-morphisms as maps of cospans in $\mathsf{A}$
%\[
%\begin{tikzpicture}[scale=1.5]
%\node (A) at (0,0.5) {$a$};
%\node (A') at (0,-0.5) {$a^\prime$};
%\node (B) at (1,0.5) {$c$};
%\node (C) at (2,0.5) {$b$};
%\node (C') at (2,-0.5) {$b^\prime$};
%\node (D) at (1,-0.5) {$c^\prime$};
%\node (E) at (3,0.5) {$d \in F(c)$};
%\node (F) at (3,-0.5) {$d^\prime \in F(c^\prime)$};
%\path[->,font=\scriptsize,>=angle 90]
%(A) edge node[above]{$$} (B)
%(C) edge node[above]{$$} (B)
%(A) edge node[left]{$f$} (A')
%(C) edge node[left]{$g$} (C')
%(A') edge node {$$} (D)
%(C') edge node {$$} (D)
%(B) edge node [left] {$h$} (D);
%\end{tikzpicture}
%\]
%together with a morphism $\iota \maps F(h)(d) \to d^\prime$ in $F(c^\prime)$.}
%\end{enumerate}
